Synthesis of Intrusion Detection Rules \\ for Increased Network Security

### 2019-06-27 paper/paper.tex
minor changes in text

\sloppy \nids{} can be categorized in two groups: 1)~rule-based and
2)~anomaly-based. A rule-based \nids{} checks if the network traffic
matches a given set of rules. A rule describes a potentially
mallicious message pattern. Figure~\Fix{small example} shows an
example rule of Suricata~\cite{suricata}, a popular open-source
\nids{}, maintained by the Open Information Security Foundation
(OISF)~\cite{oisf}. \Fix{quickly explain example} Rule-based \nids{}
focuses on known attacks whereas anomaly-based \nids{} focuses on
unanticipated issues. The hypothesis of anomaly-based \nids\ is that
attack manifestations produces traffic that deviates from regular
benign traffic that has been already observed by the network and that
deviation is noticeable. Anomaly-based \nids{}, specially those that
use machine learning, have shown useful specially for detecting denial
of service attacks~\Fix{cite cite}. In that scenario, determining the
right threshold on the rate/volume of network accesses that should
trigger an alarm is challenging. Conceptually, these two approaches
are complementary. This paper focuses on rule-based \nids{}, which are
already very popular both in industry and academia. Our analysis
focuses on Suricata, but it applies to any other rule-based \nids{}.

@@

The first statement is consequential of the inherent physical
limitation of rule-based \nids{}. The cost of checking rules is
proportional to the traffic load and the number of enabled rules. If
system capacity is insufficient to support the load, the system can
potentially miss packets and attacks\Fix{I guess we have some
  reference for this.}. To cope with that issue, system administrators
restrict the attack surface before applying \nids{}---they block
certain protocols on the firewall and configure the \nids{} to use a
selection of rules that apply to the authorized traffic. The second
statement is supported by the observation that \numundetected{} of the
attacks produced by Metasploit~\cite{metasploit}, a popular
open-source penetration testing tool, went undetected by \suri{}. For
example, \Fix{...}. (Section~\ref{sec:suri-metas-coverage} provides
more details.) Unfortunately, manually creating those rules from
scratch can be time-consuming and error-prone. Figure~\Fix{...}  shows
an example rule illustrating the complexity of the Suricata rule
language\footnote{Suricata extends the language used by
  Snort~\cite{snort}.}.  In addition, looking for rules from untrusted
sources (say, the internet) is unreliable. Collectively, these issues
suggest that synthesizing \nids{} rules is an important aid for system
administrators.

>>>>

\sloppy \nids{} can be categorized in two groups: 1)~rule-based and
2)~anomaly-based. A rule-based \nids{} checks if the network traffic
matches a given set of rules. A rule describes a potentially
mallicious message pattern. Figure~\Fix{small example} shows an
example rule of Suricata~\cite{suricata}, a popular open-source
\nids{}, maintained by the Open Information Security Foundation
(OISF)~\cite{oisf}. \Fix{quickly explain example} Rule-based \nids{}
focuses on known attacks whereas anomaly-based \nids{} focuses on
unanticipated issues. The hypothesis of anomaly-based \nids\ is that
attack manifestations produces traffic that deviates from regular
benign traffic that has been already observed by the network and that
deviation is noticeable. Anomaly-based \nids{}, specially those that
use machine learning, have shown useful specially for detecting denial
of service attacks~\cite{tutorial-ml-for-anomaly-detection-dec-2018, cordy-etal-issta19}. In that scenario, determining the
right threshold on the rate/volume of network accesses that should
trigger an alarm is challenging. Conceptually, these two approaches
are complementary. This paper focuses on rule-based \nids{}, which are
already very popular both in industry and academia. Our analysis
focuses on Suricata, but it applies to any other rule-based \nids{}.

@@

The first statement is consequential of the inherent physical
limitation of rule-based \nids{}. The cost of checking rules is
proportional to the traffic load and the number of enabled rules. If
system capacity is insufficient to support the load, the intrusion
detector can miss packets and attacks\Fix{I guess we have some
  reference for this.}. To cope with that issue, system administrators
block certain protocols on the firewall and configure the \nids{} to
use a selection of rules that apply to the authorized traffic.  The
task of selecting rules that are applicable can be non-trivial as it
depends on the set of firewall rules enabled on the network.

The second statement is supported by the observation that
\numundetected{} of the attacks produced by
Metasploit~\cite{metasploit}, a popular open-source penetration
testing tool, are undetected by \suri{}. For example,
\Fix{...}. (Section~\ref{sec:suri-metas-coverage} provides more
details.) Unfortunately, manually creating those rules from scratch
can be time-consuming and error-prone. Figure~\Fix{...}  shows an
example rule illustrating the complexity of the Suricata rule
language\footnote{Suricata extends the language used by
  Snort~\cite{snort}.}.  In addition, looking for rules from untrusted
sources (say, the internet) is unreliable. Collectively, these issues
suggest that synthesizing \nids{} rules is an important aid for system
administrators.
